---
title: "Video_Games_Case_Study"
author: "Ralph Gracia"
date: "2023-11-10"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting Up My Environment

Before starting any work with the data, I made sure I had the necessary packages installed and ready to go.

```{r loading packages}
install.packages("tidyverse")
library(tidyverse)
install.packages("rmarkdown")
library(rmarkdown)
```

## Preparing The Data

I begun my work by taking a quick look at how the data is structured. Here we can get a better understanding of the variables and values stored within the data.

To view the source of my data click here [link](https://www.kaggle.com/datasets/ibriiee/video-games-sales-dataset-2022-updated-extra-feat)

```{r preview of the data}
video_games <- read.csv("Video_Games.csv")
head(video_games)
```

## Cleaning The Data

For this case study, my goal is to find information relevant to how games have performed from 2008 to early 2017 and what the most popular genres were. In order to achieve that, the first thing I needed to address was rows that had didn't have a release year available in this dataset. I ultimately came to the conclusion that filtering out these rows wouldn't affect the overall results.

```{r filtering the data to display 2008-2017}
last_15_years <- video_games %>% filter(Year_of_Release >= 2008) %>% 
  arrange(-Global_Sales)
head(last_15_years)
```

The next step was to create new data frames for each console and see how well each console's games have performed in each region since from 2008-2017.

#### Playstation
```{r Playstation data}
ps_games <- last_15_years %>% filter(Platform =="PS" | Platform =="PS2" | Platform =="PS3" | Platform =="PSV" | Platform =="PSP"| Platform =="PS4") %>%
  filter(!is.na(Year_of_Release)) %>% arrange(-Global_Sales)
head(ps_games)
```

#### Xbox
```{r Xbox data}
xbox_games <- last_15_years %>% filter(Platform == "XB" | Platform == "X360"  | Platform == "XOne") %>%
  filter(!is.na(Year_of_Release)) %>% arrange(-Global_Sales)
head(xbox_games)
```

#### Nintendo
```{r Nintendo data}
nintendo_games <- last_15_years %>% filter(Platform =="DS" | Platform =="3DS" | Platform =="Wii" | Platform =="WiiU") %>%
  filter(!is.na(Year_of_Release)) %>% arrange(-Global_Sales)
head(nintendo_games)
```

## Visualizing Trends
My original objective is to review the data and learn what the top genres of games were on each console based on their top 100 games. After analyzing the new data frames, I used ggplot to visualize my findings. 

### Top Playstation Genres (Million USD)
```{r loading Playstation sales}
ggplot(ps_games %>% filter(Global_Sales >= 2.58), aes(Global_Sales, Genre, col=Platform)) + geom_point() + geom_jitter()
```

### Top Xbox Genres (Million USD)
```{r loading Xbox sales}
ggplot(xbox_games %>% filter(Global_Sales >= 2.34), aes(Global_Sales, Genre, col=Platform)) + geom_point() + geom_jitter()
```

### Top Nintendo Genres (Million USD)
```{r loading Nintendo Sales}
ggplot(nintendo_games %>% filter(Global_Sales >= 2.06), aes(Global_Sales, Genre, col=Platform)) + geom_point() + geom_jitter()
```

## Key Takeaways
As shown in the data visuals, the most popular genres on Playstation were: 

* Action
* Shooter 
* Sports

On Xbox:

* Action
* Shooter
* Role-Playing
* Sports

On Nintendo:

* Action
* Miscellaneous
* Platform
* Role-Playing

The following table further supports this discovery by displaying the top 50 games across all consoles. 

```{r loading top 50 games}
top_console_games <- last_15_years %>% filter(Global_Sales >= 7.38) %>%
  arrange(-Global_Sales)
head(top_console_games)
```

## The Next Step?
* Incorporate past success to future inspirations
* Learn from indie developers to evaluate popularity

My suggestion to these three companies is to take a step back and recall the formulas that allowed earlier games to succeed to begin with as well as utilize indie developers to get a better understanding of new preferences among players. 

